---
title: " "
output: github_document
---

# Exploring diabetes indicators across the US
## A county-level analysis  


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  fig.path = "README_figs/README-",
  message = FALSE,
  warning = FALSE)

Sys.getenv("CENSUS_API_KEY")

library(tidycensus)
library(tidyverse)
library(viridis)
library(sf)
library(tigris)
library(psych)
library(readxl)
library(scales)
library(cowplot)
library(jtools)
library(sjPlot)

options(tigris_class = "sf", tigris_use_cache = TRUE)

results <-  st_read("~/Desktop/DataProjects/ND and Diabetes/ArcMap/GWR/Output/GWR_RESULTS.shp")
res.HS1 <-  st_read("~/Desktop/DataProjects/ND and Diabetes/ArcMap/GWR/residuals/residual_hotspot.shp")
res.HS2 <-  st_read("~/Desktop/DataProjects/ND and Diabetes/ArcMap/GWR/residuals/residual_hotspotZOI.shp")### local R2 map 
```

## County level deprivation index

Details about the index can be found here: https://github.com/iecastro/deprivation-index.  In short, estimates are collected from the American Community Survey and a principal component analysis extracts a deprivation score for each observation.  

Deprivation index for this project was calculated using 5-year ACS estimates from 2013, at the county-level.  Higher index scores represent higher area deprivation relative to all other counties in the US.


`````````{r, echo = FALSE} 
vars <- c("B17001_002", "B17001_001", "B06009_002" , "B06009_001",
          "B09008_011", "B09008_001","B08124_002", "B08124_001", "B25014_005", 
          "B25014_006",  "B25014_007","B25014_011", "B25014_012", "B25014_013",  
          "B25014_001", "B19058_002", "B19058_001","C23002C_021", "C23002D_008", 
          "C23002C_017", "C23002D_003","B19001_002", "B19001_003", "B19001_004", 
          "B19001_005", "B19001_006", "B19001_001")

acs_us <- get_acs(geography = "county", variables = vars, year=2013, 
                  output = "wide") %>%
  mutate(pct_poverty = B17001_002E/B17001_001E,
         pct_noHS = B06009_002E / B06009_001E,
         pct_FHH = B09008_011E / B09008_001E,
         pct_mgmt = B08124_002E /  B08124_001E, 
         pct_crowd =  (B25014_005E +B25014_006E+ B25014_007E + 
                         B25014_011E + B25014_012E + B25014_013E) / B25014_001E,
         pct_pubassist = B19058_002E/B19058_001E,
         pct_unempl = (C23002C_021E + C23002D_008E)  / (C23002C_017E + C23002D_003E),
         pct_under30K =( B19001_002E+B19001_003E+B19001_004E+B19001_005E +
                           B19001_006E) / B19001_001E)

## select transformed variables
values  <-  acs_us %>% select(pct_poverty,pct_noHS,pct_FHH,pct_mgmt,pct_crowd,
                              pct_pubassist, pct_unempl,pct_under30K) %>% as.matrix()
values[is.nan(values)] <- 0
## PCA
ND <- principal(values,nfactors = 1)          
NDI_us <- cbind(acs_us,ND$scores) 
```

```{r, echo = FALSE}
NDI_us <- NDI_us %>% select(NAME,GEOID,PC1) %>% 
  separate(NAME, into = c("County","State"), sep = ",")

US_counties <- get_acs(geography = "county", variables = c("B01001_001"), 
                       output = "wide", geometry = TRUE, shift_geo = TRUE)

MapUS <- geo_join(US_counties,NDI_us, by_sp = "GEOID", by_df = "GEOID")

states <- get_acs(geography = "state", variables = c("B01001_001"), 
                  output = "wide", geometry = TRUE, shift_geo = TRUE)
```

```{r, echo=FALSE, fig.height= 8, fig.width=10}
### Map US county deprivation
ggplot() + geom_sf(data = MapUS, aes(fill = PC1)) +
  geom_sf(data=states,fill = NA,color = "#ffffff", size=.5)+
  theme_minimal() + theme(axis.text = element_blank(), legend.position = "bottom") +
  scale_fill_viridis_c(option = "inferno") +
  labs(fill = "Index", caption = "Data: US Census 5-year ACS estimates, 2013")+
  ggtitle(" ", subtitle = "County-level Deprivation")

```

## Diabetes Indicators

National data on diabetes indciators are available from the CDC (https://www.cdc.gov/diabetes/data/countydata/countydataindicators.html). Indicator estiamtes are the most up to date as of 2013 and available at the county-level only. Indicators (age-adjusted) are: 1. prevalence of diagnosed diabetes, 2. prevalence of obesity, and 3. prevalence of leisure-time physical inactivity, hereafter just inactivity. The spatial distribution of diabetes prevalence in 2013 (below) parallels that of 2007, which led to the recognition of a "diabetes belt" in southern states (https://www.scientificamerican.com/article/diabetes-belt/,https://www.ncbi.nlm.nih.gov/pubmed/21406277).  


```{r, echo = FALSE}

#### Get diabetes indicators
Diab_prev <- read_xlsx("~/Desktop/DataProjects/ND and Diabetes/Data/DM_PREV_ALL_STATES2013.xlsx")%>%
  select(State,County ,FIPS = `FIPS Codes`, DiabPrev = `age-adjusted percent`)
  
Obes_prev <- read_xlsx("~/Desktop/DataProjects/ND and Diabetes/Data/OB_PREV_ALL_STATES2013.xlsx")%>%
   select(State,County ,FIPS = `FIPS Codes`, ObPrev = `age-adjusted percent`)

Inact_prev <- read_xlsx("~/Desktop/DataProjects/ND and Diabetes/Data/LTPIA_PREV_ALL_STATES2013.xlsx") %>%
  select(State,County,FIPS = `FIPS Codes`, InactPrev = `age-adjusted percent`)

### merge datasets

merge1 <- merge(Diab_prev,Obes_prev, by.x = c("County", "State"), by.y = c("County", "State"))
merge2 <- merge(merge1,Inact_prev,by.x = c("County", "State"), by.y = c("County", "State")) %>%
  select(-c(FIPS.y))
data <- merge(NDI_us,merge2,by.x = "GEOID", by.y = "FIPS.x", all.x=TRUE) %>%
  select(-c("FIPS")) 

## merge data to spatial attributes
MapDiab <- geo_join(US_counties,data, by_sp = "GEOID", by_df = "GEOID") 
```

```{r, echo=FALSE, fig.height= 8, fig.width=10}
### map diabetes prevalence 
### map diabetes prevalence 
ggplot() + geom_sf(data=MapDiab, aes(fill = as.numeric(DiabPrev)/100)) +
  geom_sf(data=states,fill = NA,color = "#ffffff", size=.5) +
  scale_fill_viridis_c(option = "magma",labels = scales::percent) +
  theme_minimal() + theme(axis.text = element_blank(), legend.position = "bottom") +
  labs(fill = "Prevalence ", caption = "Data: CDC County-level Indicators, 2013")+
  ggtitle(" ", subtitle = "County-level, age-adjusted prevalence of diagnosed Diabetes in adults")
```

## Data Analysis

Obesity and physical inactivity are well-known risk factors for developing diabetes.  Therefore, in order to assess whether area deprivation is associated with diabetes prevalence, independent of obesity and inactivity, a blocked multiple regression was fitted.  First we estimated the effects of obesity and inactivity; afterwards, deprivation variable was introduced.

```{r, echo = FALSE}
#### data cleaning 
data <- data %>% select(-c(State.x,County.y)) %>% 
  as.tibble()

data$DiabPrev <- as.numeric(data$DiabPrev)
data$ObPrev <- as.numeric(data$ObPrev)
data$InactPrev <- as.numeric(data$InactPrev)

## blocked model 
lm1 <- lm(DiabPrev~ObPrev+InactPrev, data = data)
lm2 <- lm(DiabPrev~ObPrev+InactPrev + PC1, data = data)

```

### Regression results

```{r, echo = FALSE}
## compare models
sjt.lm(lm1,lm2,depvar.labels = c("Block 1","Block 2"),
       pred.labels = c("Obesity Prevelence", "Inactivity Prevalence","Area Deprivation"),
       show.r2 = TRUE, show.fstat = TRUE, show.ci = TRUE,show.se = TRUE)
```



#### Increases in county-level deprivation predict increases in diabetes prevalence.

```{r, echo = FALSE}
## plots 
ggplot(data, aes(PC1,DiabPrev)) + geom_point() +
  stat_smooth() + theme_classic() +
  labs(x = "County-level Deprivation Index", y = "Diagnosed Diabetes Prevalence (%)")
```


It's possible that area deprivation is a moderator in the association between obesity and diabetes.

```{r, echo = FALSE}
ggplot(data, aes(PC1,DiabPrev)) + geom_point(aes(color = ObPrev/100)) +
  stat_smooth() + theme_classic() +
  labs(x = "County-level Deprivation Index", y = "Diagnosed Diabetes Prevalence (%)",
       color = "Obesity") +
  scale_color_viridis_c(option="inferno", direction = -1, labels = scales::percent)
```

### Interaction Term

Results summary 

```{r, echo=FALSE}
## interaction model
int <- lm(DiabPrev~InactPrev+ObPrev*PC1, data = data)
summ(int,center = TRUE)
```


```{r, echo=FALSE}
interact_plot(int, pred = ObPrev, modx = PC1,
              interval = TRUE, int.type = "confidence", 
              x.label = "Obesity Prevalence (%)", y.label = "Diagnosed Diabetes Prevalence (%)") +
             theme_apa(legend.pos = "topleft") + 
              ggtitle("County-level Deprivation", 
              subtitle = "moderates the relationship between obesity and diabetes")
```


## Spatial Analysis

All spatial analysis was conducted in ArcMap (version 10.2). A geographically weighted regression (GWR) was fitted with an adapative kernel and AICCc bandwith parameter. 

GWR estimates local models across an area by allowing the relationship between variables to vary by location. Inactivity variable was excluded from this model due to multicollinearity with obesity variable. 

### Results 

The GWR used 404 neighbors for each local estimation and accounted for 80% of variability in diabetes outcome. Local models accounted between 20% to 77% of variability in the outcome.

GWR summary:

 Neighbors           : 404  
 ResidualSquares     : 3022.5957759565522  
 EffectiveNumber     : 83.492913855248489   
 Sigma               : 0.99411194002147207  
 AICc                : 8921.5582814951449  
 R2                  : 0.80040748447829668  
 R2Adjusted          : 0.79502414949644495  



```{r, echo=FALSE, fig.height= 9, fig.width=11}
 ggplot() + geom_sf(data=results, aes(fill = LocalR2)) +
  geom_sf(data=states,fill = NA,color = "#ffffff", size=.5)+
  theme_minimal() + theme(axis.text = element_blank()) +
  scale_fill_distiller(type = "seq", direction = 1, labels = scales::percent) +
  labs(fill="Explained Variance") +
  ggtitle("Spatial variability in goodness-of-fit", subtitle = "determined by Local R-squared values")
```


### Spatial variability of effects

#### Obesity prevelance

```{r, echo=FALSE, fig.height= 9, fig.width=11}

ggplot() + geom_sf(data=results, aes(fill = C1_obesity)) +
  geom_sf(data=states,fill = NA,color = "#ffffff", size=.5)+
  theme_minimal() + theme(axis.text = element_blank()) +
  scale_fill_viridis_c(option = "cividis", direction = -1) +
  labs(fill = expression(paste("Estimate", "(",beta,")"))) +
  ggtitle("Effect of obesity on diabetes prevalence")

ggplot() + geom_sf(data=results, aes(fill = StdErrC1_o)) +
  geom_sf(data=states,fill = NA,color = "#ffffff", size=.5)+
  theme_minimal() + theme(axis.text = element_blank()) +
  scale_fill_viridis_c(option = "magma", direction = 1) +
  labs(fill = "Std. Error") +
  ggtitle("", subtitle="Precision of estimates")
```

#### Area deprivation

```{r, echo=FALSE, fig.height= 9, fig.width=11}

ggplot() + geom_sf(data=results, aes(fill =C2_PC1)) +
  geom_sf(data=states,fill = NA,color = "#ffffff", size=.5)+
  theme_minimal() + theme(axis.text = element_blank()) +
  scale_fill_viridis_c(option = "cividis", direction = -1) +
  labs(fill = expression(paste("Estimate", "(",beta,")"))) +
  ggtitle("Effect of deprivation on diabetes prevalence")

ggplot() + geom_sf(data=results, aes(fill = StdErrC2_P)) +
  geom_sf(data=states,fill = NA,color = "#ffffff", size=.5)+
  theme_minimal() + theme(axis.text = element_blank()) +
  scale_fill_viridis_c(option = "magma", direction = 1) +
  labs(fill = "Std. Error") +
  ggtitle("", subtitle="Precision of estimates")

```


### Standardized residuals from GWR model


```{r, echo=FALSE,fig.height= 9, fig.width=11}
###### residual maps
## std residuals 

ggplot() + geom_sf(data=results, aes(fill = StdResid)) +
  geom_sf(data=states,fill = NA,color = "#000000", size=.5)+
  theme_minimal() + theme(axis.text = element_blank()) +
  scale_fill_distiller(type = "div", palette = "RdYlBu") 
```


```{r, echo=FALSE, fig.height= 9, fig.width=11}
## hot spot inversed distance
ggplot() + geom_sf(data=res.HS1, aes(fill = Gi_Bin)) +
  geom_sf(data=states,fill = NA,color = "#000000", size=.5)+
  theme_minimal() + theme(axis.text = element_blank()) +
  scale_fill_distiller(type = "div", palette = "RdYlBu",breaks = c(-3,-2,0,2,3), labels =  c("99%","95%", "Not Sig.", "95%", "99%")) + 
  labs(fill = "Confidence") +
  ggtitle("Hot Spot analysis of residual values", subtitle = "with an Inverse Distance spatial relationship")
```

```{r, echo=FALSE, fig.height= 9, fig.width=11}                     
## hot spot ZOI
ggplot() + geom_sf(data=res.HS2, aes(fill = Gi_Bin)) +
  geom_sf(data=states,fill = NA,color = "#000000", size=.5)+
  theme_minimal() + theme(axis.text = element_blank()) +
  scale_fill_distiller(type = "div", palette = "RdYlBu",breaks = c(-3,-2,0,2,3), 
                       labels =  c("99%","95%", "Not Sig.", "95%", "99%"))  + 
  labs(fill = "Confidence") +
  ggtitle("Hot Spot analysis of residual values", 
          subtitle = "with a Zone of Indifference spatial relationship")
```


## Limitations



  
